{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2021-11-28 pheno.tex
Update of 8.1-8.3

The main MSSM contributions are expected to come from chargino--stop\footnote{We usually expect higher generation off-diagonal soft terms to be larger due to RGE running controlled by Yukawa couplings.} and charged Higgs--top loops, as shown in Figs.~\ref{btosdiags} c) and d), respectively. However,  there is little room for effects from supersymmetry since the current experimental world average is ${\rm Br}(B\to X_s\gamma)=(3.32\pm 0.15)\cdot 10^{-4}$ (PDG 2010). This means that either the charged Higgs is heavy enough and the stop-scharm soft mass term small enough not to contribute significantly, or that there are cancellations between the two contributions. This is both an example of a process that puts significant constraints on off-diagonal soft terms, and a process that is important in constraining the possible values of the charged Higgs mass.

%%%
\subsection{$B_s \to \mu^+\mu^-$}
%%%
The process  $B_s \to \mu^+\mu^-$ is another FCNC process as either the bottom or the strange quark must change flavour in order to couple to the muons. The Standard Model process is shown in Fig.~\ref{bsmudiags} a), where the CKM factor in the $W$ vertex allows one of the quarks to change flavour, but also suppresses the decay rate. For strange historical reasons involving darts and illegal substances this process is known as a {\bf penguin diagram}.

On top of the small CKM factor, the process also suffers from what is called {\bf helicity suppression} in the Standard Model. The $Z$-boson is spin-1, while the starting point meson $B_s$ is spin-0 (pseudoscalar), meaning that the spins of the quarks are opposite. At some point in the diagram the helicity (chirality) must ``flip". This introduces an extra suppression proportional to $m_\mu^2/M_{B_s}^2$, making the expected rate extremely small and sensitive to supersymmetry contributions. We get a similarly suppressed process for $B_d$ with a $\bar d$-quark instead of the $\bar s$ in the initial state.

@@

The predicted SM branching ratios for these processes are~\cite{Bobeth:2013uxa}:
\begin{eqnarray}
{\rm Br}(B_s \to \mu^+\mu^-) &=& (3.65 \pm 0.23)\cdot 10^{-9},\\
{\rm Br}(B_d \to \mu^+\mu^-) &=& (1.06 \pm 0.09)\cdot 10^{-10}.
\end{eqnarray}
First evidence for the $B_s$ decay was shown by the LHCb collaboration in 2012. The final observation required combining Run I data from both LHCb and CMS, and was published in 2014~\cite{CMS:2014xfa}. The current values are:
\begin{eqnarray}
{\rm Br}(B_s \to \mu^+\mu^-) &=&  2.8^{+0.7}_{-0.6}\cdot 10^{-9},\\
{\rm Br}(B_d \to \mu^+\mu^-) &=&  3.9^{+1.6}_{-1.4}\cdot 10^{-10},
\end{eqnarray}
where one should keep in mind that the $B_d$ decay has only evidence at $3.2\sigma$ significance.

In the MSSM there are contributions from process such as shown in Fig.~\ref{bsmudiags} b). These contributions are proportional to $\tan^6\beta$, which makes the decay process highly sensitive to scenarios with large $\tan\beta$. To see this dependence, notice that $\mu$ couples to the mediating heavy higgses $H/A^0$ through the Yukawa term $y^l_{22} L_2 H_d \overline{E}_2$ in the superpotential, and the Yukawa constant in this term, $y^l_{22} = y_\mu$, is connected to the fermion mass through $m_\mu = y_\mu v\cos\beta$. Thus this vertex is proportional to $1/\cos\beta$ or $\tan\beta$, giving a factor $\tan^2\beta$ in the amplitude squared.\footnote{Remember that in the limit of large $\tan\beta$
\begin{equation}
\cos\beta= \pm \frac{1}{\sqrt{1+\tan^2\beta}}= \pm \frac{1}{\tan\beta\sqrt{1+\frac{1}{\tan^2\beta}}}\simeq\pm \frac{1}{\tan\beta}.
\end{equation}
}

Furthermore, a chargino(higgsino)--stop loop can couple the strange and bottom quarks to the higgs. These couplings are proportional to the bottom Yukawa coupling $y_b$, from the superpotential terms $y_{33}^dQ_3H_d\bar D_3$, which appears in the stop--chargino--bottom vertex, and the  $y_{32}^uQ_3H_d\bar D_2$, which appears in the strange--chargino--stop vertex. Both these Yukawa couplings are proportional to $y_b$ and thus to  $1/\cos\beta$, giving a further factor of $\tan^4\beta$ in the amplitude squared. This $\tan\beta$ dependence makes $B_s \to \mu^+\mu^-$ an excellent channel for  discovering supersymmetry, and puts very stringent bounds on the sparticle masses in large $\tan\beta$ scenarios.

>>>>

The main MSSM contributions are expected to come from the chargino--stop\footnote{We usually expect higher generation off-diagonal soft terms to be larger due to RGE running controlled by Yukawa couplings.} and charged Higgs--top loops, as shown in Figs.~\ref{btosdiags} c) and d), respectively. However,  there is little room for effects from supersymmetry since the current experimental world average is ${\rm Br}(B\to X_s\gamma)=(3.32\pm 0.15)\cdot 10^{-4}$~\cite{HFLAV:2019otj}. This means that either the charged Higgs is heavy enough and the stop-scharm soft mass term small enough not to contribute significantly, or that there are cancellations between the two contributions. This is both an example of a process that puts significant constraints on off-diagonal soft terms, and a process that is important in constraining the possible values of the charged Higgs mass.

%%%
\subsection{$B_s \to \mu^+\mu^-$}
%%%
The process  $B_s \to \mu^+\mu^-$ is another FCNC process as either the bottom or the strange quark must change flavour in order to get a coupling to the muons. The Standard Model process is shown in Fig.~\ref{bsmudiags} a), where the CKM factor in the $W$ vertex allows one of the quarks to change flavour, but also suppresses the decay rate. For strange historical reasons involving darts and illegal substances this process is known as a {\bf penguin diagram}~\cite{Shifman:1995hc}.

On top of the small CKM factor, the process also suffers from what is called {\bf helicity suppression} in the Standard Model. The $Z$-boson is spin-1, while the starting point meson $B_s$ is spin-0 (pseudoscalar), meaning that the spins of the quarks in the meson are opposite. At some point in the diagram the helicity (chirality) must ``flip". This introduces an extra suppression proportional to $m_\mu^2/M_{B_s}^2$, making the expected rate extremely small and sensitive to supersymmetry contributions. We get a similarly suppressed process for $B_d$ with a $\bar d$-quark instead of the $\bar s$ in the initial state.

@@

The predicted SM branching ratios for these processes are~\cite{Bobeth:2013uxa}:
\begin{eqnarray}
{\rm Br}(B_s \to \mu^+\mu^-) &=& (3.65 \pm 0.23)\cdot 10^{-9},\\
{\rm Br}(B_d \to \mu^+\mu^-) &=& (1.06 \pm 0.09)\cdot 10^{-10}.
\end{eqnarray}
First evidence for the $B_s\to\mu^+\mu^-$ decay was shown by the LHCb Collaboration in 2012. The current value for the branching ratio from combined measurements at LHCb, ATLAS and CMS is~\cite{LHCb:2020zud}
\begin{eqnarray}
{\rm Br}(B_s \to \mu^+\mu^-) &=&  2.69^{+0.37}_{-0.35}\cdot 10^{-9},\\
{\rm Br}(B_d \to \mu^+\mu^-) &<&  1.6\cdot 10^{-10},
\end{eqnarray}
where the value for $B_d$ is an upper limit at the 90\% confidence level. It is interesting to observe that the measured value is lower than the Standard Model prediction at around the $2\sigma$ level, this has lead to speculations of destructive interference from new physics.

If we limit ourselves to scenarios with only diagonal soft terms the dominant contribution in the MSSM is from process such as shown in Fig.~\ref{bsmudiags} b), with a chargino and up-type squark loop. These contributions are proportional to $\tan^6\beta$, which makes the decay process highly sensitive to scenarios with large $\tan\beta$. To see this dependence, notice that $\mu$ couples to the mediating heavy higgses $H/A^0$ through the Yukawa term $y^l_{22} L_2 H_d \overline{E}_2$ in the superpotential, and the Yukawa constant in this term, $y^l_{22} = y_\mu$, is connected to the fermion mass through $m_\mu = y_\mu v\cos\beta$. Thus this vertex is proportional to $1/\cos\beta$ or $\tan\beta$, giving a factor $\tan^2\beta$ in the amplitude squared.\footnote{Remember that in the limit of large $\tan\beta$
\begin{equation}
\cos\beta= \pm \frac{1}{\sqrt{1+\tan^2\beta}}= \pm \frac{1}{\tan\beta\sqrt{1+\frac{1}{\tan^2\beta}}}\simeq\pm \frac{1}{\tan\beta}.
\end{equation}
}

Furthermore, a chargino(higgsino)--stop loop can couple the strange and bottom quarks to the Higgs. These couplings are proportional to the bottom Yukawa coupling $y_b$, from the superpotential terms $y_{33}^dQ_3H_d\bar D_3$, which appears in the stop--chargino--bottom vertex, and the  $y_{32}^dQ_3H_d\bar D_2$, which appears in the strange--chargino--stop vertex. Both these Yukawa couplings are proportional to $y_b$ and thus to  $1/\cos\beta$, giving a further factor of $\tan^4\beta$ in the amplitude squared. This $\tan\beta$ dependence makes $B_s \to \mu^+\mu^-$ an excellent channel for  discovering supersymmetry, and puts very stringent bounds on the sparticle masses in large $\tan\beta$ scenarios.
